---
title: "Dr. Martin's R Notebook by Kyle (Zhuqing) Wang"
output: html_notebook
---

This is Dr. Martin's R Notebook by Kyle.

Install packages before load libraries.

*install.packages("gapminder")*
*install.packages("tidyverse")*

load libraries up front.

```{r}
library(gapminder)
library(tidyverse)
```

creating a gapminder data frame

```{r}
data("gapminder")
gapminder
```

The first command does not do anything, the data showed up when we run the second command.

Glance at the gapminder dataframe using *summary()*

```{r}
summary(gapminder)
```

a histogram of the *populations* column (log-transformed)

```{r}
attach(gapminder) # attach() the gapminder dataframe to the environment so that never have to type those $ symbols again. without this attach function, all of the below operations on the columns of dataframe "gapminder" would have to be written as "gapminder$pop", rather than "pop", or we would have to specify "data = gapminder" in various functions.
hist(log(pop))
```

scatterplot of *life expectancy* (y-axis) vs. log-transformed *gdp-per-capital* using the base plot() function

```{r}
plot(lifeExp ~ gdpPercap)
```

more linear with log()

```{r}
plot(lifeExp ~ log(gdpPercap))
```

a t-test of life-expectancy differences between South Africa and Ireland

```{r}
df1 <- gapminder %>% 
  select(country, lifeExp) %>% 
  filter(country == "South Africa" | # | means or;
           country == "Ireland")
t.test(data = df1, lifeExp ~ country)
```

color-enhanced ggplot single-plot of life expectancy vs. gdp-per-capita, colored by continent, and dot-size by population, with the smooth function

```{r}
gapminder %>% 
  filter(gdpPercap < 50000) %>% 
  ggplot(aes(x=log(gdpPercap), y=lifeExp, col=continent, size=pop)) +
  geom_point(alpha=0.3) +
  geom_smooth(method=lm)
```

faceted ggplot scatterplot, colored by year, faceted by continent, with the smooth function

```{r}
gapminder %>% 
  filter(gdpPercap < 50000) %>% 
  ggplot(aes(x=log(gdpPercap), y=lifeExp, col=year, size=pop)) +
  geom_point(alpha=0.3) +
  geom_smooth(method=lm) +
  facet_wrap(~continent)
```

a linear model test of life expectancy by gdp-per-capita and population, with the summary displayed

```{r}
summary(lm(lifeExp ~ log(gdpPercap)+log(pop), data=gapminder))
```